Bird beak's elaboration and exploration

### 2021-09-20 Analysis/02.1-MCMCglmm_mini_chains_cluster.Rmd
Cluster script

All the custom code and custom functions called in this document can be downloaded and installed in `R` as the standalone working package `beer` (Beak Elaboration and Exploration in R).
Although `beer` is intentionally designed to be portable and shareable, we advise workers to use it with moderation and tailor its consumption to their specific research needs.
Some analyses in `beer` do take some time to run so we advise you do not operate heavy machinery while using it.
Don't use it while driving.
The following code snippets are used to illustrate the implementation of the functions.
See <!-- §§§run_replicates_script --> for running the actual analyses.

@@

# MCMCglmm mini chains analyses

To estimate the variance-covariance matrix for the phylogeny and each clade (see [01-Data_preparation.Rmd](01-Data_preparation.Rmd)) we run multivariate generalised linear mixed models (MCMCglmm) using the `MCMCglmm` package [@MCMCglmm].

## Model

We ran a general multi-response model with `§§§N_traits` of the trait space as fixed effects (using a Gaussian distribution) and and the residual covariance structure.
We used `§§§N_clades + 1` random effects (one for the whole phylogeny and one for each clade).
<!-- % NC: I know this is in the data prep, but can you specify an example MCMCglmm model here to make it clearer what you mean? -->

## Mini chains

To increase the speed of the analysis and take phylogenetic uncertainty into account, we used a "mini chains" approach.
It runs multiple MCMCglmm analyses on multiple trees and pulls the results together into one larger MCMCglmm that contains more variation due to phylogenetic uncertainty (similar to @mulTree - ADD LINK?).
The size of the chains are optimised for speed. <!-- % NC: Editted as multree is not familiar to everyone, so keeping it general and linking to multree is better -->

### Mini-chains parametrization

We first ran three independent MCMC chains with the model and data described above using the consensus tree and flat priors.
We ran these models (for `§§§N_generations` sampling every `§§§N_nitt` with no burnin) until the three chains reached an effective sample size >> 200 (§§§and other diagnosis). 
<!-- % NC: I thought in MCMCglmm the nitt parameter was total number of iterations, and thin gives how often the chain is sampled? I think you've called nitt generations here but better to be consistent with MCMCglmm?
 -->

@@

We set up a mini-chain to be a MCMCglmm running with 1) the model described above, 2) a random tree, and 3) the estimated priors (with a degree of parameter belief of 5% `nu = 0.05`) to run for `nitt` generations were `nitt` is equal to the previously estimated burnin phase + 100 sampled iterations (`nitt = burnin + thin * 100`).
Each mini-chain thus provided us with 100 posterior data points.
The two main advantages of this mini-chain approach is that 1) they are much faster to run since no diagnosis of convergence is necessary and the chains are only run for a relatively short time (which allow several chains to crash/fail without losing all the outputs); and 2) they take into account tree uncertainty without having to run the complete MCMCglmm on all trees (c.f. @mulTree).

@@

We then run the MCMCglmm over $N_trees$ to get an estimate of the variance-covariance matrices including phylogenetic uncertainty.

```{r, running_replicates, eval = FALSE}
## The number of replicates
n_reps <- 10

## Run the required chains
results <- run.mini.chains(mini_chains, replicates = n_reps, path = "../Data/Processed/MCMC/")

## Combining the results as one big MCMC
combined_results <- combine.mini.chains(results)

## Saving the results
save(combined_results, file = "../Data/processed/combined_results.rda")
```

>>>>

This script is a working script to run the chains on the SHARC Sheffield cluster.
Refer to `02-MCMCglmm_mini_chains.Rmd` for more details and reproducibility

@@

# All birds shapespace analyses

## Mini-chains parametrization

@@

@@

The chains are then run on the cluster
TODO: details

### Combining the model data

```{r}
## Preparing the model files
model_list <- list()
class(model_list) <- c("beer", "mini.chains")
models_path <- "../Data/Cluster/allbirds_8748t_8Dshapespace_31levels/cluster_out/"
## Loading all models
for(file in list.files(models_path)) {
    load(paste0(models_path, file, collapse = "/"))
    model_list[[length(model_list)+1]] <- model[[1]]
}

## Combining the model files
allbirds_8748t_8Dshapespace_31levels <- combine.mini.chains(model_list)

## Saving the results
save(allbirds_8748t_8Dshapespace_31levels, file = "../Data/processed/allbirds_8748t_8Dshapespace_31levels.rda")
```
